---
title: "Reef P2 - Report"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
date: '2022-05-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#libraries
library(tidyr)
library(dplyr)
library(tidyverse)
library(maps)
library(ggplot2)
library(sjPlot)
library(cvTools)
library(ozmaps)
library(vcdExtra)
library(corrplot)
library(RColorBrewer)
library(caret)
```

```{r read_data, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
# Reading data
setwd('..')
new_reef = read_csv("./Data/lr_version_merged_mean.csv")

# Cleaning names
data = new_reef %>% janitor::clean_names()

## removing NA data entries
data <- na.omit(data)

## changing date to date
# data$date <- as.Date(data$date, format = "%d-%b-%y")

# Convert bleached
data$bleached <- ifelse(data$average_bleaching > 0, "1", "0")
# 0 = not bleached
# 1 = bleached
data %>% count(bleached)

# Convert to numeric
data$bleached <- as.factor(data$bleached)
```

# Executive Summary

Coral reefs are the most diverse marine ecosystems in the world. They provide billions of dollars of economic value through Coastal protection, food, tourism, and medicine from the sea. Coral bleaching leads to coral morbidity and mortality, which leads to coral loss. Global atmosphere-ocean circulation models predict that few coral reefs will survive in the ocean for the next 80 years.

We know that warmer ocean temperatures linked to climate change are the primary factors in coral bleaching. In the interdisciplinary project with the MARS students, we identified an interesting correlation between ocean currents, thermal activity, and its effect on coral. However, upon further research, conflicting literature states that, while ocean currents can transport excess nutrients that make corals more susceptible to bleaching (DeCarlo et al., 2020), stronger currents can also flush out toxins aiding in the process of coral recovery (Nakamura et al., 2003).

Our team looked at the relationship between average ocean current velocity and the effect it has on the bleaching of corals in the Great Barrier Reef. In doing so, we can confirm our hypothesis that ocean currents do have a significant impact on coral bleaching. The motivation for our group to implement the project is that it enables marine scientists to predict and visualize the likelihood of coral bleaching based on specific parameters of the Great Barrier Reef. Our overall aim was to implement different classification models on Python Dash to enable marine scientists to predict and visualize the chance of coral bleaching given specific parameters within the Great Barrier Reef. This will allow further attention on areas where corals are more susceptible to stressing factors. 

Figure 1

# Method

```
- Communicate the results to a general scientist, note this is different to the target audience of your product. 
- A clear description of your innovation and approach to the scientific problem.
- Include all relevant data summaries only.
- What is your approach and which tools are used and why? 
- What are the key findings? and use figures to illustrate the results.
- How robust and generalizable is your finding or your product?
- Main code and technical details of your approach in the appendix.
```
The following models were used for variable selection and classification: logistic regression (LR), random forest (RF), and support vector machine (SVM).

We analyzed these model accuracies, combined with feature selection to determine which classifier and features were best for our goals. By comparing the model performances, we concluded the random forest classifier was the most accurate for classifying coral bleaching. Our models also showed which of the characteristic variables were more significant in determining the main factors that lead to coral bleaching. 

### Data Collection 

```
Clear description of the approach in data collection, developed model, the evaluation strategies from a data-science perspective. Here we refer to all types of evaluation metrics including graphical, qualitative and quantitative metric.

Reef project - Clear description of the approach in data collection to address appropriate research questions, developed model, the evaluation strategies from a marine science perspective,
```
The data used in this project is merged from the DATA3888 Reef Check data as well as ocean current netCDF files taken from AIMS (Australian Institute of Marine Science). The ocean data was collated from the years 2010 to 2015 by month. We used `mean_cur` (average of current velocity) as our main variable from ocean currents. Most of the fluctuations of the tidal currents are averaged out, thus allowing some of the large-scale current flow patterns to be seen.

**Data Collection**

When merging data sets, we found that latitude and longitude cannot be accurately matched. Using past data and Jackson's guidance, the team chose to divide the latitude and longitude of each observation into a range rather than an exact value. At the same time, two methods were adopted when merging the data sets, thus obtaining two data sets. A small amount of data is suitable for logistic regression, and a large amount of data is suitable for machine learning, which can help the team get a more accurate model.

**Data Prepossessing**

Our main variable of interest was `Average_bleaching`. We discovered there were more 0 values than any other values. For our classification task, our team chose to convert this variable to a binary variable with a threshold of 0. In this case, where the average bleaching equals 0, it would be classified as "bleached", and where average bleaching is greater than 0, it would be "not bleached". However, in our regression model, it would be reversed (average bleaching equals 0 showed no signs of bleaching, and greater than 0 would imply there has been previous bleaching.

### Models

We tested our merged data on 3 classification models - Logistic Regression, Support Vector Machine and Random Forest. With average bleaching as the response model, we used the same variables for our 3 models. This included the variables:

- ClimSST (Climatological Sea Surface Temperature)
- Temperature_Kelvin (temperature in kelvin)
- Temperature_Kelvin_Standard_Deviation
- SSTA_Frequency (Sea Surface Temperature Anomaly)
- SSTA_Frequency_Standard_Deviation
- TSA_Frequency_Standard_Deviation
- mean_cur (mean ocean current velocity)

The team chose an algorithm that combines efficiency and accuracy, meaning a balance between extremely short runtime and high accuracy. In the ML classification model below, SVM and Random Forest are from SKLearn's suite of ML algorithms. Logistic regression comes from the glm function of the R language. To further reduce equipment impact, time measurements were made on the same computer.

The accuracy of each model is determined by cross-fold validation on the training set, while the speed of the classifier is simultaneously determined by setting a timer for the prediction stage.
The models our team investigated are as follows:

### Logistic Regression Model

We used a Logistic Regression Model to predict the outcome of bleaching against our chosen predictor variables from our merged data. Our logistic regression model works by returning an estimate of the probability of coral bleaching occurring between 0 and 1. 

First, by looking at the significance of our chosen variables, we want to see whether there is a strong correlation between average current velocity and coral bleaching. We decided on p-values less than 0.05 determining a more significant relationship to bleaching. From our model, the mean current velocity had a value of 0.89. This indicates that ocean velocity had no effect in the event of coral bleaching. However, the SSTA frequency (number of times that the sea surface temperature anomaly had a change of more than one degree Celsius) and the temperature variables showed significance associated to the outcome. 

The coefficient estimate of `mean_cur` is a positive 0.25, which indicates that a higher average current velocity is associated the an increased probability that the reef will be susceptible to coral bleaching. This alludes to Mcclanahan et al., (2005), that more intense bleaching was found on sites with higher currents and bleaching was observed to be positively associated with higher water flow. Likewise, the coefficients for `clim_sst` and `temperature_kelvin` were 0.01, suggesting higher temperatures were positively associated to coral bleaching. Since it is difficult to distinguish between water flow, temperature, and the other environmental factors of the ocean, it is harder to conclude the main factors responsible for bleaching.


### Support Vector Machine

In the SVM model,  the accuracy of the model is 83%. Through the linear kernel function, the coefficient between bleaching and current velocity is -1.89e to the power of 5, which indicates that there is negatively correlated between them, although they are correlated but too weak，so we can say there is no significance between coral bleaching and  ocean currents. 


### Random Forest

The random forest model is composed of multiple decision trees, and each decision tree is used to judge whether or not to bleach. Using multiple decision trees can make the judgment more comprehensive and accurate. A decision tree can be understood as a logical process for you to judge whether reef are bleached. However, one decision tree is more susceptible to outliers, and overfitting is also more likely to occur. Multiple people make decisions together, which will produce a fairer judgment. So random forest can reduce the impact of outliers, reducing the possibility of overfitting. In the random forest model, the accuracy of the model is 87%. From the importance of the features in the model, the effect of ocean currents velocity on whether coral reefs are bleached is small. The importance of mean cur is only 0.005. This means that our hypothesis is rejected in this model, there is no relationship between ocean current velocity and coral bleaching. 

### Evaluation strategies

SVM: Accuracy: 0.83
Interpretability: the coefficient between bleaching and current velocity is -0.0000189
no significance in coral bleaching related to ocean current velocity. 
Scalability: <1 second per 10%, 50%, 70% and 100% sample of data.

Figure 2


# Results

```
A clear justification of the final approach based on the proposed evaluation strategies. Ensuring multiple evaluation strategies are used.

A clear description of the deployment process. An engaging and clear illustration of the product (games, shiny app, learning device etc) with a discussion of concepts from multiple disciplines.
```
There shows to be no correlation between average ocean velocity and the event of coral bleaching in our Great Barrier Reef. All 3 models reflected the same results. The model evaluations showed our Random Forest model had the most accuracy in classifying whether coral bleaching would occur with from our chosen variables. With an overall accuracy of 87%, we have decided to implement this model as part of our final product. 


python dash stuff..

### Innovation

blahlafafbla

# Discussion

```
Discussion of potential shortcomings or issues associated with the development process or the product with reference to both disciplines. Identification of future work or improvement in both disciplines. Conclusion adequately summarises the project and identification of future work.
```

### Limitations

A limitation in our method and data collection was that we only looked at the data on corals of the Great Barrier Reef. This could have an impact on our hypothesis that ocean currents would influence coral bleaching as our data is focused on the Great Barrier Reef. Whereas currents on bleaching on a more global scale could show different results. Furthermore, the scope of our project only covered average ocean velocity and not the direction of the current. The MARS students have identified that, even among the Great Barrier Reef, many different currents flow in and around the reef, also fluctuating upon season. The current direction has also been a considered factor in causing our reefs to be more susceptible to bleaching. These issues can be addressed in our future works.

Another limitation of our project was that we only tested our merged data on 3 classification models. Random Forest had the most accuracy in classifying whether coral bleaching would occur from our chosen variables. With an overall accuracy of 87%, however, we cannot guarantee that the Random Forest model is the best fitting model.

```
ocean currents can transport excess nutrients that make corals more susceptible to bleaching , stronger currents can also flush out toxins aiding in the process of coral recovery . But we only assume a relationship between average ocean current velocity and the effect it has on the bleaching of corals. In the Great Barrier Reef, we can control the ocean currents within a certain range. When the collected ocean currents exceed a certain value, we can discard that part of the ocean currents data, and perhaps get a more accurate conclusion.


What is the last part trying to say?
```

### Future work

On a bigger project, we could address our limitations of using only a small proportion of our data. To check whether ocean currents have an impact on coral bleaching, we could compare our reef datasets to global ocean current data. Although the power to run analyses will exponentially increase since global netCDF files are significantly larger, this will include more data for us to draw more meaningful conclusions. Also, studying the direction of ocean currents can help identify other factors in coral bleaching. As ocean currents connect different water sources, this can also influence the temperature of the water flowing in from the equator and subtropics, etc. 

# Conclusion

# Appendix

### Figure 1 {.tabset}

Here, we have graphed an initial look at our data.

#### Coral bleaching
```{r, warning = FALSE, message = FALSE}
world <- map_data("world")
oz_states <- ozmap_states

ggplot(oz_states) + geom_sf() + coord_sf() +
  geom_point(
    data = data,
    aes(reef_longitude, reef_latitude, colour = bleached),
    alpha = 0.7) + ggtitle("Amount of coral points bleached vs not bleached in the GBF") + 
  ylab("Latitude") + xlab("Longitude") + scale_colour_manual(labels = c("not bleached", "bleached"), values = c("red", "blue"))
```

#### Average coral bleaching
```{r}
ggplot(oz_states) + geom_sf() + coord_sf() +
  geom_point(
    data = data,
    aes(reef_longitude, reef_latitude, size = average_bleaching, colour = average_bleaching),
    alpha = 0.7) + ggtitle("Average coral bleaching along the GBF") + 
  ylab("Latitude") + xlab("Longitude") 

```


# Student contributions  
(50-200 words)

# References

- https://www.barrierreef.org/the-reef/threats/coral-bleaching 
- https://oceanservice.noaa.gov/facts/coral_bleach.html 
- DeCarlo ...
- Nakamura ...

```
# Presentation
- Correct grammar
- All figures and tables captioned, axes, headings, legends
- Only show necessary figures (Appendix)
```

