---
title: "Regression"
output: html_document
date: '2022-05-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(tidyverse)

library(sjPlot)
library(cvTools)

library(vcdExtra)
library(corrplot)
library(RColorBrewer)
```

Reading in the reef data:
```{r}
reef = read_csv("Data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv")

new_reef = read_csv("Data/lr_version_merged_mean.csv")

# Check what the reef dataset looks like
head(new_reef)
```
Data cleaning: 

- Working on the copy of the reef dataset called data
- Cleaned names so it would be easier to work with the variables

```{r}
## clean variable names
data = new_reef %>% janitor::clean_names()

names(data)
```
- Removing NA rows, we cannot work with missing values that we cannot access from the dataset
- Converting format of entries in the variables we can use
```{r}
## removing NA data entries
data <- na.omit(data)

## changing date to date
# data$date <- as.Date(data$date, format = "%d-%b-%y")

## changing temperature to celcius
data$temperature_celcius <- data$temperature_kelvin - 273.15

## longitude and latitude maximum and minimum to match dataset
```

Assumption: For the logistic regression, the predictors are meant to be binary. So we convert the entries of average_bleaching >0 to be 1.
```{r}
# We have to sort out bleached and not bleached corals
data$bleached <- ifelse(data$average_bleaching > 0, "1", "0")
# 0 = not bleached
# 1 = bleached
data %>% count(bleached)

# Convert to numeric
data$bleached <- as.numeric(data$bleached)

```

Assumption: Observations from the dataset can be related to each other?

Assumption: Correlation 


Fitting the Logistic Regression:

```{r}
glm = glm(bleached ~ clim_sst + temperature_kelvin + temperature_kelvin_standard_deviation + ssta_frequency + ssta_frequency_standard_deviation + tsa_frequency_standard_deviation + mean_cur, family = binomial, data = data)
summary(glm)

glm
```
Predicting accuracy

accuracy
- RMSE by CV

stability
- how different model performs across all models
- spread of rmse (boxplot)
spread across cv
  
scalability
- how well it scales across different sizes
- time code in r? how long models take to run (10%, 20, 50, 100)
how steep the line is tells u how scalable the model is

compare the models

interpretability
- something u need to say about the model

