---
title: "Merging the eReefs data"
author: "Jackson Zhou"
output:
  html_document:
    theme: united
    highlight: kate
---

<style>
h1, h2, h3, h4, h5 {text-align: left;}
body {text-align: justify;}
.leaflet {margin: auto;}
</style>

Loading in packages:

```{r Loading in packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(sf)
library(nngeo)
```

Loading in the required data:

```{r Loading in the data}
reef <- read_csv("./Data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv",
                 show_col_types = FALSE) %>% 
  clean_names()

currents <- read_csv("./Data/eReefs-aggregated-monthly-data.csv",
                     show_col_types = FALSE) %>% 
  clean_names()
```

We filter the reef data to only include Australian (in particular, Queensland) data from 2010 - 2017 (the range of the ocean currents data).
In addition, we add a numeric month column to match that of the currents data, and an index column so that processing the merging results is easier.

```{r Filtering the reef data}
reef_filtered <- reef %>% 
  filter(year >= 2010 & year <= 2017) %>% 
  filter(country == "Australia") %>% 
  filter(state_province_island == "Queensland") %>% 
  mutate(month = match(sapply(str_split(date, "-"), "[[", 2), month.abb),
         index = row_number())
```

Checking the geographic range covered by both datasets:

```{r Checking the geographic ranges}
cat("Reef data range:", range(reef_filtered$latitude_degrees), "|", range(reef_filtered$longitude_degrees))
cat("\nCurrents data range:", range(currents$latitude), "|", range(currents$longitude))
```

The currents data is seen to cover the reef data, which is good.
First, we perform a spatial cross-join on the filtered reef data with the currents data.
After that, we average the rows which are joined across each index.

```{r Spatial cross-join and further merging}
reef_filtered_sf <- reef_filtered %>% 
  st_as_sf(coords = c("longitude_degrees", "latitude_degrees"), remove = FALSE)

currents_sf <- currents %>% 
  st_as_sf(coords = c("longitude", "latitude"), remove = FALSE)

# k nearest neighbours merge
reef_crossed <- st_join(reef_filtered_sf, currents_sf, join = st_nn, maxdist = 0.5, k = 88) %>% 
  as.data.frame() %>% 
  arrange(index, year.y, month.y) %>% 
  select(-geometry)

reef_summarised <-  reef_crossed %>% 
  group_by(index) %>% 
  summarise(eta = mean(eta),
            salt = mean(salt),
            temp = mean(temp),
            wspeed_u = mean(wspeed_u),
            wspeed_v = mean(wspeed_v),
            mean_wspeed = mean(mean_wspeed),
            u = mean(u),
            v = mean(v),
            mean_cur = mean(mean_cur))

reef_merged <- left_join(reef_filtered, reef_summarised, by = "index")
```

Fitting a basic logistic regression model to the data:

```{r Fitting a basic model}
reef_merged_processed <- reef_merged %>% mutate(bleached = as.factor(ifelse(average_bleaching > 0, 1, 0)))

glm(bleached ~ eta + salt + temp + mean_cur + mean_wspeed, 
    family = binomial(link = "logit"), 
    data = reef_merged_processed) %>% 
  summary()
```
